The Magnetorotational Instability Prefers Three Dimensions

### 2019-08-22 paper/paper.tex
Added phi vs SSC to figure 1, use mean field alpha plot instead

Numerical simulations aimed at understanding the MRI in accretion disks consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}. 
These ``channel modes'' are exact non-linear solutions for $x$-shearing-periodic domains, and can only saturate via parasitic shear instabilities \citep{1994ApJ...432..213G}.
We use impenetrable, stress-free boundary conditions that are more applicable to stellar interiors. 
But even in disks, any kind of finite radial extent will cutoff the unbounded growth of channel modes. 
We reconcile our results with earlier simulations by showing that the MRI indeed prefers two dimensions for the larger values of criticality found in disk simulations. 
Figure~\ref{fig:phi} shows the phase angle $\phi = \arctan k_y/k_z$ of the fastest growing mode as function of $\SSC$.
Above $\SSC \gtrsim 2.05$ $\phi=0$, indicating that axisymmetric modes have the fastest growth rates.
For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC \simeq 4.84$ (most works use similar values).
Thus, we predict axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{phi_vs_ssc_grid.pdf}
  \caption{Angle $\phi$ of mode with respect to $z$ axis vs $\SSC$. The inset shows the transition from axisymmetric modes to three dimensional modes as $\SSC$ decreases below $2.05$. The shaded region indicates stability to all perturbations; the stability limit for three dimensional modes is $\SSC \simeq 0.102$.}
  \label{fig:phi}
\end{figure}

>>>>

Numerical simulations aimed at understanding the MRI in accretion disks consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}. 
These ``channel modes'' are exact non-linear solutions for $x$-shearing-periodic domains, and can only saturate via parasitic shear instabilities \citep{1994ApJ...432..213G}.
We use impenetrable, stress-free boundary conditions that are more applicable to stellar interiors. 
But even in disks, any kind of finite radial extent will cutoff the unbounded growth of channel modes. 
We reconcile our results with earlier simulations by showing that the MRI indeed prefers two dimensions for the larger values of criticality found in disk simulations. 
Figure~\ref{fig:phi} shows the phase angle $\phi = \arctan k_y/k_z$ of the fastest growing mode as function of $\SSC$.
Above $\SSC \gtrsim 2.05$ $\phi=0$, indicating that axisymmetric modes have the fastest growth rates.
For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC \simeq 4.84$ (most works use similar values).
Thus, we predict axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{{mean_field_alpha_run_11_output.h5}.pdf}
  \caption{Mean field $\alpha$ coefficient.}
  \label{fig:phi}
\end{figure}
